### 2016-03-11 docs/ms_communitySynchrony.Rmd
updates supp table order

Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out five communities (Fig. 2).
Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S2).
Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
Removing competition caused, at most, a 5% change in synchrony (Table S2).
Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2), with an average 6% increase over synchrony from IBM simulations on a five meter area.
The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig. 3).
For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
Results for synchrony of percent cover are qualitatively similar (Figs. S2-S3) but more variable, as expected by theory [@Loreau2008a].

@@

However, our results do contrast with a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives species synchrony in grasslands [@Gross2014].
The apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
Gross et al.'s results are based on rapidly assembling experimental communities that are, at most, 11 years old.
The natural, relatively undisturbed communities we studied consist of species that have co-occurred for decades [@Chu2015] and represent a much later stage of community assembly.
In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments the processes driving competitive exclusion are still operating.
Such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
In contrast, synchrony may be driven by environmental fluctuations in older communities. In other words, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct repeat the analysis of @Gross2014 every few years.

CONCLUSIONS
-----------
Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2).
A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
Such a conclusion is only half correct.
The perennial grasses we studied do have similar responses to the environment (Table X1-X10), which will tend to synchronize dynamics.
However, if community-wide species synchrony is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
In agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
Such a conclusion is in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
Recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
Future research on the influence of community assembly on synchrony could reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.

>>>>

Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out five communities (Fig. 2).
Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S1).
Overall, species' temporal random effects in the statistical vital rate models are positively correlated (Table S2).
Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
Removing competition caused, at most, a 5% change in synchrony (Table S2).
Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2), with an average 6% increase over synchrony from IBM simulations on a five meter area.
The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig. 3).
For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
Results for synchrony of percent cover are qualitatively similar (Figs. S2-S3) but more variable, as expected by theory [@Loreau2008a].

@@

However, our results do contrast with a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives species synchrony in grasslands [@Gross2014].
The apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
Gross et al.'s results are based on rapidly assembling experimental communities that are, at most, 11 years old.
The natural, relatively undisturbed communities we studied consist of species that have co-occurred for decades [@Chu2015] and represent a much later stage of community assembly.
In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments the processes driving competitive exclusion are still operating.
Such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
In contrast, synchrony may be driven by environmental fluctuations in older communities. In other words, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.

CONCLUSIONS
-----------
Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
When we removed environmental variation, we found that synchrony decreased in four out of the five grassland communities we studied (Fig. 2).
A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
Such a conclusion is only half correct.
The perennial grasses we studied do have similar responses to the environment (Table S2), which will tend to synchronize dynamics.
However, if community-wide species synchrony is less than 1, as it is in all our focal communities, some degree of compensatory dynamics must be present [@Loreau2008a].
In agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.
Such a conclusion is in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
Recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
Future research on the influence of community assembly on synchrony could reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.

### 2016-03-11 docs/online_supporting_information.Rmd
updates supp table order

Tables
------
```{r avg_spp_interactions, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)
####
####  Get Vital Rate Satistical Results Files ----------------------------------
####
vital_rates <- c("growth", "surv", "recruit")
result_files <- list.files("../results/")
vr_files <-result_files[grep(paste(vital_rates,collapse="|"), 
                              list.files("../results/"))]

@@

```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)

>>>>

Tables
------
```{r perc_diff_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)

```{r corr_yeareffs, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)
####
####  Read in regression fits --------------------------------------------------
####
all_files <- list.files("../results/")
param_files <- all_files[grep("param*", all_files)]
num_vitals <- length(param_files)

list_yr_effects <- list()
for(j in 1:num_vitals){
  param_list <- readRDS(paste0("../results/",param_files[j]))
  vital_name <- unlist(strsplit(param_files[j], "_"))[1]

  if(vital_name != "recruit"){

    site_names <- names(param_list)
    site_year_list <- list()
    for(do_site in site_names){
      tmp_site <- param_list[[do_site]]
      num_spp <- length(tmp_site) 
      tmp_yr_int <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
      tmp_yr_slope <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
      for(i in 1:num_spp){
        spp_df <- tmp_site[[i]]
        tmp_yr_int[,i] <- spp_df$Intercept.yr
        col2use <- grep("logarea", colnames(spp_df))
        col2use2 <- grep(".yr", colnames(spp_df[col2use]))
        tmp_yr_slope[,i] <- spp_df[,col2use[col2use2]]
      } # end species within site loop
      site_year_list[[do_site]][["intercept"]] <- tmp_yr_int
      site_year_list[[do_site]][["slope"]] <- tmp_yr_slope
    } # end site within vital rate loop

  } # end NOT recruit loops

  if(vital_name == "recruit"){

    site_names <- names(param_list)
    site_year_list <- list()
    for(do_site in site_names){
      tmp_site <- param_list[[do_site]]
      tmp_yrs <- tmp_site[grep("intcpt.yr", rownames(tmp_site)),"Mean"]
      stringlist <- strsplit(names(tmp_yrs), ",")
      num_spp <- length(unique(sapply(stringlist, "[[", 2)))
      mat_yrs <- t(matrix(tmp_yrs, nrow=num_spp, byrow = TRUE))
      site_year_list[[do_site]] <- mat_yrs
    } # end site within vital rate loop

  } # end YES recruit loops

  list_yr_effects[[vital_name]] <- site_year_list
} # end vital rate loop

# ## Plot histograms of year effects by vital rate
# vital_rates <- names(list_yr_effects)
# par(mfrow=c(3,5))
# for(do_vital in vital_rates){
#   tmp_vital <- list_yr_effects[[do_vital]]
#   site_names <- names(tmp_vital)
#   for(do_site in site_names){
#     tmp_site <- tmp_vital[[do_site]]
#     hist(tmp_site, main=paste(do_site,do_vital), xlab="Year Effect", xlim=c(-3,8))
#   }
# }

##  Get standard deviation of year effect posterior

####
####  Calculate average correlation of year effects ----------------------------
####
vital_rates <- names(list_yr_effects)
out_df <- data.frame(site=NA, vital_rate=NA, term=NA, corr=NA)
for(do_vital in vital_rates){
  tmp_vital <- list_yr_effects[[do_vital]]
  site_names <- names(tmp_vital)
  for(do_site in site_names){
    tmp_site <- tmp_vital[[do_site]]

    if(do_vital != "recruit"){
      for(i in 1:length(tmp_site)){
        tmp_cor <- cor(tmp_site[[i]])
        avg_cor <- mean(tmp_cor[upper.tri(tmp_cor)])
        tmp_out <- data.frame(site=do_site, vital_rate=do_vital, term=names(tmp_site)[i], corr=avg_cor)
        out_df <- rbind(out_df, tmp_out)
      }
    }

    if(do_vital == "recruit"){
      tmp_cor <- cor(tmp_site)
      avg_cor <- mean(tmp_cor[upper.tri(tmp_cor)])
      tmp_out <- data.frame(site=do_site, vital_rate=do_vital, term="intercept", corr=avg_cor)
      out_df <- rbind(out_df, tmp_out)
    }

  }
}
cor_df <- out_df[2:nrow(out_df),]
cor_cast <- dcast(cor_df, site+term~vital_rate)

cor.cap <- "Correlations of species' year random effects for each site by term, where term refers to the random effect on the slope or the intercept."
print(xtable(cor_cast, caption = cor.cap),
  caption.placement="top",
  include.rownames = FALSE, 
  sanitize.colnames.function = identity,
  comment=FALSE)
```

```{r avg_spp_interactions, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(ggplot2)
library(xtable)
####
####  Get Vital Rate Satistical Results Files ----------------------------------
####
vital_rates <- c("growth", "surv", "recruit")
result_files <- list.files("../results/")
vr_files <-result_files[grep(paste(vital_rates,collapse="|"), 
                              list.files("../results/"))]

@@
